---
title: "A Tale of Three Penguins"
subtitle: "Exploring the Palmer Penguins Dataset Through Scrollytelling"
format: 
  closeread-html:
    cr-style:
      section-background-color: white
---

```{r}
#| include: false
library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(tidyr)

# Set theme for all plots
theme_set(theme_minimal(base_size = 14))
```

The Palmer Archipelago in Antarctica is home to three remarkable penguin species. Let's explore their story through data.

:::{.cr-section}

In the harsh Antarctic environment, three penguin species have made their home on the Palmer Archipelago islands. @cr-intro-map

:::{#cr-intro-map}
```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 6

# Create a visualization of penguin counts by species
penguins %>%
  filter(!is.na(species)) %>%
  count(species, island) %>%
  mutate(island = paste(island, "Island")) %>%
  ggplot(aes(x = species, y = n, fill = species)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~island) +
  scale_fill_manual(values = c("Adelie" = "#fb7504", 
                                "Chinstrap" = "#c65bcf", 
                                "Gentoo" = "#067476")) +
  labs(title = "Penguin Species Distribution Across Islands",
       x = "Species",
       y = "Number of Penguins Observed") +
  theme(plot.title = element_text(size = 18, face = "bold"))
```
:::

:::{#cr-intro-counts}

```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 6
#| 
penguins %>%
  filter(!is.na(species)) %>%
  count(species, island) %>%
  mutate(island = paste(island, "Island")) %>%
  ggplot(aes(x = species, y = n, fill = species)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 5) +
  facet_wrap(~island) +
  scale_fill_manual(values = c("Adelie" = "#fb7504", 
                                "Chinstrap" = "#c65bcf", 
                                "Gentoo" = "#067476")) +
  labs(title = "Penguin Species Distribution Across Islands",
       x = "Species",
       y = "Number of Penguins Observed") +
  theme(plot.title = element_text(size = 18, face = "bold"))
```
:::
The dataset contains observations of 344 penguins, carefully measured and documented by researchers. @cr-intro-counts

Each penguin was measured across several physical characteristics. [@cr-intro-counts]{highlight="344"}

## The Three Species

Let's meet our Antarctic residents, starting with the most widespread species. @cr-species-plot

:::{#cr-species-plot}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

# Create detailed species comparison
species_summary <- penguins %>%
  filter(!is.na(body_mass_g), !is.na(flipper_length_mm)) %>%
  group_by(species) %>%
  summarise(
    avg_mass = mean(body_mass_g),
    avg_flipper = mean(flipper_length_mm),
    .groups = "drop"
  )

penguins %>%
  filter(!is.na(body_mass_g), !is.na(flipper_length_mm)) %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_manual(values = c("Adelie" = "#fb7504", 
                                 "Chinstrap" = "#c65bcf", 
                                 "Gentoo" = "#067476")) +
  labs(title = "Body Mass vs Flipper Length by Species",
       x = "Flipper Length (mm)",
       y = "Body Mass (g)",
       color = "Species") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 20, face = "bold"))
```
:::

[**Adelie penguins**]{style="color: #fb7504;"} are the smallest of the three, recognizable by their distinctive white eye rings. [@cr-species-plot]{highlight="Adelie"}

[**Chinstrap penguins**]{style="color: #c65bcf;"} get their name from the thin black line running under their chin. [@cr-species-plot]{highlight="Chinstrap"}

[**Gentoo penguins**]{style="color: #067476;"} are the largest, with bright orange-red bills. [@cr-species-plot]{pan-to="-50%,20%" scale-by="1.5" highlight="Gentoo"}

Notice how the species cluster distinctly in size and flipper length. [@cr-species-plot]

## Sexual Dimorphism

[**Male**]{style="color: #0891b2;"} penguins tend to be larger than [**female**]{style="color: #d946ef;"} penguins across all three species. @cr-dimorphism

:::{#cr-dimorphism}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

penguins %>%
  filter(!is.na(sex), !is.na(body_mass_g), !is.na(bill_length_mm)) %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = sex)) +
  geom_point(size = 3, alpha = 0.6) +
  facet_wrap(~species, scales = "free") +
  scale_color_manual(values = c("female" = "#d946ef", "male" = "#0891b2")) +
  labs(title = "Sexual Dimorphism: Bill Length vs Body Mass",
       x = "Bill Length (mm)",
       y = "Body Mass (g)",
       color = "Sex") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 18, face = "bold"),
        strip.text = element_text(size = 14, face = "bold"))
```
:::

This size difference is most pronounced in the **Gentoo** penguins. [@cr-dimorphism]{pan-to="-72%,0%" scale-by="1.2"}

Despite size differences, [**males**]{style="color: #0891b2;"} and [**females**]{style="color: #d946ef;"} follow similar trends in body proportions. @cr-dimorphism

## Bill Dimensions

The shape of a penguin's bill tells us about their feeding behavior and ecological niche. @cr-bills

:::{#cr-bills}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

penguins %>%
  filter(!is.na(bill_length_mm), !is.na(bill_depth_mm)) %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color = species, shape = species)) +
  geom_point(size = 4, alpha = 0.7) +
  scale_color_manual(values = c("Adelie" = "#fb7504", 
                                 "Chinstrap" = "#c65bcf", 
                                 "Gentoo" = "#067476")) +
  labs(title = "Bill Dimensions: A Species Signature",
       x = "Bill Length (mm)",
       y = "Bill Depth (mm)",
       color = "Species",
       shape = "Species") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 20, face = "bold"))
```
:::

[**Adelie penguins**]{style="color: #fb7504;"} have shorter, deeper bills - perfect for their diet. [@cr-bills]{pan-to="30%,30%" scale-by="1.2"}

[**Chinstrap penguins**]{style="color: #c65bcf;"} have intermediate bill dimensions. [@cr-bills]{pan-to="-20%,30%" scale-by="1.2"}

[**Gentoo penguins**]{style="color: #067476;"} possess longer, shallower bills adapted to their feeding strategy. [@cr-bills]{pan-to="-20%,-40%" scale-by="1.2"}

## Temporal Patterns

:::{#cr-empty}
:::

The timing of nesting varies across species and islands. @cr-empty

:::{#cr-temporal-blank}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

penguins %>%
  filter(!is.na(year)) %>%
  count(year, species) %>%
  ggplot(aes(x = year, y = n, fill = species)) +
  scale_fill_manual(values = c("Adelie" = "#fb7504", 
                                "Chinstrap" = "#c65bcf", 
                                "Gentoo" = "#067476")) +
  labs(title = "Penguin Observations Over Time",
       x = "Year",
       y = "Number of Observations",
       fill = "Species") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 20, face = "bold")) +
  scale_x_continuous(breaks = c(2007, 2008, 2009))
```
:::

:::{#cr-temporal}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

penguins %>%
  filter(!is.na(year)) %>%
  count(year, species) %>%
  mutate(alpha_value = if_else(species == "Adelie", 1, 0.7)) %>%
  ggplot(aes(x = year, y = n, fill = species, alpha = alpha_value)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("Adelie" = "#fb7504", 
                                "Chinstrap" = "#c65bcf", 
                                "Gentoo" = "#067476")) +
  scale_alpha_identity() +
  labs(title = "Penguin Observations Over Time",
       x = "Year",
       y = "Number of Observations",
       fill = "Species") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 20, face = "bold")) +
  scale_x_continuous(breaks = c(2007, 2008, 2009))
```
:::

:::{#cr-temporal-adelie}
```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 7

penguins %>%
  filter(!is.na(year)) %>%
  count(year, species) %>%
  mutate(n = if_else(species == "Adelie", n, 0)) %>%
  ggplot(aes(x = year, y = n, fill = species)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("Adelie" = "#fb7504", 
                                "Chinstrap" = "#c65bcf", 
                                "Gentoo" = "#067476")) +
  labs(title = "Penguin Observations Over Time",
       x = "Year",
       y = "Number of Observations",
       fill = "Species") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 20, face = "bold")) +
  scale_x_continuous(breaks = c(2007, 2008, 2009))
```
:::

Data were collected over three breeding seasons from 2007 to 2009. @cr-temporal-blank

[**Adelie penguins**]{style="color: #fb7504;"} were observed consistently across all years. @cr-temporal-adelie

The study captured population dynamics during a critical period. @cr-temporal


:::

---

## Data Source

Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network. The dataset was prepared by Allison Horst, Alison Hill, and Kristen Gorman.

**Citation:** Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/
